---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# abstractify

<!-- badges: start -->
<!-- badges: end -->

Have you ever been tinkering in R and realize after writing some block of code
"Oh this should really be a helper function at this point" ?
You then begin the process of wrapping `foo <- function(...) { ... }`
around your code, then you try to run `foo` only to find that you forgot to
include something in your function arguments that your block of code needed?
This addin allows you to highlight a block of code, run `abstractify()` from
the command palette, and the block will be wrapped with a new function 
definition *but with all arguments that need to be specified*.

## Installation

You can install the development version of abstractify like so:

``` r
# Either or
devtools::install_github("tsostarics/abstractify")
remotes::install_github("tsostarics/abstractify")
```

## Example

Usage is hard to show in just text, I'll need to make one of those terminal
recording gifs later.
But the basic idea is illustrated below. 
Say we've been playing around with `mtcars` and want to filter by the number
of cyls, count the number of remaining cars, then add 2.
We might have an R script with the following lines:

```{r example}
library(abstractify)
library(dplyr)

necessary_cyls <- c(3, 4)
fudge_factor <- 2

mtcars |> 
  dplyr::filter(cyl %in% necessary_cyls) |> 
  dplyr::summarize(n = n() + fudge_factor)
```

At this point we might want to make a function that does the filtering
and summarizing for some other dataset that we have, parameterized by
arbitrary values of `necessary_cyls` and `fudge_factor` which currently exist
just as global variables in our R session.
We can highlight the last three lines and run `abstractify()` to get the
following automatically:

```r
new_function <- function(mtcars,
                         necessary_cyls,
                         fudge_factor) {
  mtcars |> 
  dplyr::filter(cyl %in% necessary_cyls) |> 
  dplyr::summarize(n = n() + fudge_factor)
}
```

If we instead wanted `fudge_factor` to stay constant in the helper function,
then we would highlight from `fudge_factor <- 2` to `...+ fudge_factor)` and
run `abstractify()` in the same way:

```r
new_function <- function(mtcars,
                         necessary_cyls) {
  fudge_factor <- 2
  
  mtcars |>
    dplyr::filter(cyl %in% necessary_cyls) |> 
  dplyr::summarize(n = n() + fudge_factor)
}
```

Now the helper is only parameterized in terms of `mtcars` and `necessary_cyls`,
which you can rename within the scope of the helper function by highlighting
and using the existing Rename Symbol In Scope command 
(default: `Ctrl+Shift+Alt+M`).
Finally, again since it's hard to show just in text, after running `abstractify()`
the function name (`new_function`) will be highlighted so that you can
immediately start writing the actual desired function name.

### Usage notes/assumptions

I made this because I frequently find myself needing to turn scratchwork
into helper functions for my research (e.g., creating a big ggplot, then needing
to make the same plot with different datasets).
As such, `abstractify()` (currently) makes a few assumptions when identifying
variables to abstract over:
`abstractify()` will abstract over any symbols it finds so long as they 
*exist in the global environment*.
The following things won't work perfectly & i'll try to extend or fix things
later.

 - Running `abstractify()` on a codeblock within a function will miss objects
 that exist only in the scope of the function
 - Instances where you would expect an error to be thrown because an object
 can't be found will likely result in `abstractify()` not catching it. This has
 the benefit of skipping over something like `cyl` in `filter(mtcars, cyl==3)`
 but has the downside of skipping over `foo` in `filter(mtcars, cyl==foo)` if
 you never ran `foo <- 3` in your R session.
 - An error is thrown if using `abstractify()` in a qmd file due to a known
 error with the rstudioapi when running the `reformatCode` command in files
 other than `.R` files.
